Parallel Simulation of Adaptive Random Boolean Networks

### 2016-07-14 Paper/paper.tex
Cosmetics, Knuth algo

	\paragraph{Knuth algorithm.} The algorithm is based on the following observation: $\mathbf{S}(\tau) = \mathbf{S}(2\tau)$ if and only if $\tau$ is a multiple of an attractor length $\mu$, i. e. $\tau = n\mu$. In order to find $\tau$, two copies of a Boolean network are launched, one copy performs one state update per iteration, and the other one performes two. Then these states are compared. Nodes' mean state on an attractor, which is required to apply ADRR, can be found without determining the value of $\mu$. First, one should find sum of nodes' states on $t \in (\tau, 2\tau]$ by calulating state sum vector $\Sigma = \mathbf{S}(2\tau) + \mathbf{S}(2\tau + 1) - \mathbf{S}(\tau)$ on each iteration. After $\tau$ is found, mean state of the $i$-th node on the attractor is $\langle\sigma_i\rangle = \Sigma_i / \tau$. Having found $\tau$ it is possible to obtain the length of the attractor by finding $\mu$ such that $\mathbf{S}(\tau) = \mathbf{S}(\tau + \mu)$.

@@

	\begin{algorithm}[ht!]
		\label{algo:gpu}
		GenerateNetwork()\;
		\While{ epoch $<$ epochsCount }
		{
			$gpuNetwork \leftarrow$ ConvertNetworkToGpuRepresentation()\;
			ZeroInitialize($stateSum$)\;
			CopyToGpu($gpuNetwork$, $currentState$, $stateSum$)\;
			\While{attractor not found}
			{ 	
				UpdateStateOnGpu($currentState$, $stateSum$)\;
			}
			CopyFromGpu($stateSum$)\;
			$nodesToRewire \leftarrow$ ChooseRandomNodes($N_{rew}$)\;
			ActivityDependentRewiring($nodesToRewire$, $stateSum$)\;
			epoch = epoch + 1\;
		}
		\caption{GPGPU algorithm for ARBN simulation}
	\end{algorithm}

	\paragraph{} The performance of the GPU implementation highly depends on the Boolean network representation it uses. Since a node with inputs \(\mathbf{N_i}\) has \(2^{|\mathbf{N_i}|}\) possible combinations of inputs' state, it is a common approach to store a Boolean function of the node as \(1 \times 2^{|\mathbf{N_i}|}\) vector \(\mathbf{b_i}\). We can represent these vectors as \(N \times 2^{\max{|\mathbf{N_i}|}}\) matrix \(B\).

>>>>

	\paragraph{Knuth algorithm.} The algorithm is based on the following observation: $\mathbf{S}(\tau) = \mathbf{S}(2\tau)$ if and only if $\tau$ is a multiple of an attractor length $\mu$, i. e. $\tau = n\mu$. In order to find $\tau$, two copies of a Boolean network are launched, one copy performs one state update per iteration, and the other one performes two. Then the states $\mathbf{S}(t)$ and $\mathbf{S}(2t)$ are compared, and $\tau$ is considered found if they are equal. To apply ADRR, one needs to know nodes' mean state on an attractor. It can be found without determining the value of $\mu$. First, one should find sum of nodes' states on $t \in (\tau, 2\tau]$ by calulating a state sum vector $\Sigma(t) = \Sigma(t - 1) + \mathbf{S}(2t) + \mathbf{S}(2t + 1) - \mathbf{S}(t)$ on each iteration. After $\tau$ is found, mean state of the $i$-th node on the attractor is $\langle\sigma_i\rangle = \Sigma_i(\tau) / \tau$. Having found $\tau$ it is possible to obtain the length of the attractor by finding $\mu$ such that $\mathbf{S}(\tau) = \mathbf{S}(\tau + \mu)$.

@@

	\begin{algorithm}[ht!]
		\label{algo:gpu}
		GenerateNetwork()\;
		\While{ epoch $<$ epochsCount }
		{
			$gpuNetwork \leftarrow$ ConvertNetworkToGpuRepresentation()\;
			ZeroInitialize($stateSum$)\;
			CopyToGpu($gpuNetwork$, $currentState$, $stateSum$)\;
			\While{attractor not found}
			{ 	
				UpdateStateOnGpu($currentState$, $stateSum$)\;
			}
			CopyFromGpu($stateSum$)\;
			$nodesToRewire \leftarrow$ ChooseRandomNodes($N_{rew}$)\;
			ActivityDependentRewiring($nodesToRewire$, $stateSum$)\;
			$epoch \leftarrow  epoch + 1$\;
		}
		\caption{GPGPU algorithm for ARBN simulation}
	\end{algorithm}

	The performance of the GPU implementation highly depends on the Boolean network representation it uses. Since a node with inputs \(\mathbf{N_i}\) has \(2^{|\mathbf{N_i}|}\) possible combinations of inputs' state, it is a common approach to store a Boolean function of the node as \(1 \times 2^{|\mathbf{N_i}|}\) vector \(\mathbf{b_i}\). We can represent these vectors as \(N \times 2^{\max{|\mathbf{N_i}|}}\) matrix \(B\).
